---
title: "Untitled"
author: "Nirmala Dhakal"
date: "10/12/2021"
output: html_document
---

#linear Regression:

##correaltion in r 

```{r}
library(palmerpenguins)
cor(penguins$bill_length_mm,
    penguins$body_mass_g, 
    use = "pairwise.complete")  #[1] 0.5951098 
##this is correlation between bill length and body mass.

```


#find the Pearson correlation bill length and bill depth for Gentoo penguins.

```{r}
library(palmerpenguins)
gentoo <- penguins |> filter(species == "Gentoo")
cor.test(penguins$bill_length_mm,
    penguins$body_mass_g, 
    use = "pairwise.complete") #pearson's product mooment correlation
```

#OR we can use this chunk

```{r}
penguins |> 
  filter(species == "Gentoo") |> 
  drop_na(bill_length_mm, bill_depth_mm) |> 
  summarise(corr = cor(bill_length_mm, bill_depth_mm))
```

#Find the confidence intervals of this correlation. Is it significant?


```{r}
ct <- cor.test(gentoo$bill_length_mm, gentoo$bill_depth_mm, use = "complete.obs")
ct 
#95% confidence interval.it is significant.there is a 95% chance that the confidence interval of [0.5262952 0.7365271] contains the true population correlation coefficient between bill length and bill depth in gentoo penguins.

#95 percent confidence interval does not include 0, means it is significant,further away from zero either negatively or positively,it is more significant.
```

#calculate r^2 and how much variation in data is explained by correlation.

```{r}
(0.6433839)^2 =0.413 #41% variance is explained in data. R^2 always increase with more no. of predictor.
```


#Fitting a least-squares model in R

```{r}
library(palmerpenguins)
gentoo <- penguins |> filter(species == "Gentoo")
mod <- lm(bill_length_mm ~ body_mass_g, data = gentoo)
mod
```


#We want to test how bill length varies with body mass in Chinstrap penguins

#make plot to show data

```{r}
chinstrap <- penguins |> 
  filter(species == "Chinstrap")
ggplot(chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point() +
  labs(x = "Body mass g", y = "Bill length mm")

```

#fit a suitable model to the data

```{r}
mod <- lm(bill_length_mm ~ body_mass_g, data = chinstrap)
mod
summary(mod)#Coefficients:
             #Estimate Std. Error t value Pr(>|t|)    
             #(Intercept) 3.217e+01  3.443e+00   9.344 1.07e-13 ***
              #body_mass_g 4.463e-03  9.176e-04   4.863 7.48e-06 ***
#std error is uncertainty in data.
#t avlue and p value very far from 0, so result is significant.
```

```{r}
av <- car::Anova(mod)#anova can be used to calculate F statistics. if F value large means large no. of DF.#Large f value also shows significance.
av

```


#Examine the diagnostic plots. Is everything OK?

```{r diagnostics, fig.height = 8}
library(ggfortify)
autoplot(mod, which = 1:6)
```



## Categorical predictors

#We want to test whether Adelie penguins differ in body mass between islands.

# Make a suitable plot to show the raw data

```{r}
library(ggbeeswarm)
adelie <- penguins |> filter(species == "Adelie")

island_mass_plot <- ggplot(adelie, aes(x = island, y = body_mass_g)) +
  geom_violin() +
  geom_quasirandom() + # like geom_sina but more evenly spaced
  labs(x = "Island", y = "Body mass g")
island_mass_plot

```

- Fit a suitable model to the data

```{r}
mod2 <- lm(body_mass_g ~ island, data = adelie)
coef(mod2) #(Intercept)     islandDream islandTorgersen 
           3709.659091      -21.266234       -3.286542 

           #(mean mass in bisco)(diff in mass betw bisco and dream)(diff in mass beween bisco and torgersen)
           
           
```



- Interpret the model coefficients

```{r}
coef(mod2) 
```

- Is the model statistically significant?

```{r}
av2 <- car::Anova(mod2)
av2 #P VALUE greater than 0.05, not significant

```

- Include the predicted mass with confidence interval on the plot

```{r}

nd <- distinct(adelie, island)
pred_mass <- augment(mod2, newdata = nd, interval = "confidence")
island_mass_plot +
  geom_pointrange(aes(y = .fitted, ymin = .lower, ymax = .upper), data = pred_mass, colour = "red", size = 1.5)
  

```

- Do a post-hoc test to test for differences between pairs of islands (in real analysis only do this if overall model is significant) (Hint - use `multcomp` package but DO NOT use `library(multcomp)` - it also loads `plyr` and causes conflicts)
- plot the result of the post-hoc test and interpret the results


